---
title: "analyze_dragonfly_wings"
author: "Jacob Idec"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hypotheses

Test

## Q1: What are the most informative model formulas predicting wing pigments and pattern PCs using climate, flight style, sex, and body size without interactions?

First, step model and get VIFs with all variables

```{r vifs_and_step}

# black
vifsAndStep(wings_black,response="black",formula="flight_type + wing_type + Sex + 
            temp_indv + temp_mean_sp + temp_sd_sp + temp_q5_sp + temp_q95_sp + 
            se_indv + se_mean_sp + prec_indv + prec_mean_sp
            + fore_length",mixed=FALSE)
vifsAndStep(wings_black,response="black",formula="flight_type + wing_type + Sex + temp_sd_sp + 
    temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length",mixed=FALSE)

# brown
vifsAndStep(wings_brown,response="brown",formula="flight_type + wing_type + Sex + 
            temp_indv + temp_mean_sp + temp_sd_sp + temp_q5_sp + temp_q95_sp + 
            se_indv + se_mean_sp + prec_indv + prec_mean_sp
            + fore_length",mixed=FALSE)
vifsAndStep(wings_brown,response="brown",formula="flight_type + wing_type + Sex + temp_sd_sp + 
    temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length",mixed=FALSE)

# yellow
vifsAndStep(wings_brown,response="yellow",formula="flight_type + wing_type + Sex + 
            temp_indv + temp_mean_sp + temp_sd_sp + temp_q5_sp + temp_q95_sp + 
            se_indv + se_mean_sp + prec_indv + prec_mean_sp
            + fore_length",mixed=FALSE)
vifsAndStep(wings_brown,response="yellow",formula="flight_type + wing_type + Sex + temp_sd_sp + 
    temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length",mixed=FALSE)

# black and brown
vifsAndStep(wings_bb,response="bb_total",formula="flight_type + wing_type + Sex + 
            temp_indv + temp_mean_sp + temp_sd_sp + temp_q5_sp + temp_q95_sp + 
            se_indv + se_mean_sp + prec_indv + prec_mean_sp
            + fore_length",mixed=FALSE)
vifsAndStep(wings_bb,response="bb_total",formula="flight_type + Sex + temp_indv + temp_q5_sp + 
    temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length",mixed=FALSE)

```

Formula for black: is "flight_type + wing_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length"

Formula for brown: "flight_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + fore_length"

Formula for yellow: "wing_type + Sex + temp_indv + temp_mean_sp + se_mean_sp"

The fact that we can predict the amount of black better than the amount of black and brown suggests that black is somehow biologically different from brown

Run lmers with these:

```{r cars}

getPlotLM(wings_black,model_type = "lmer",response="black",formula="flight_type + wing_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length + (1 | species)")

getPlotLM(wings_brown,model_type = "lmer",response="brown",formula= "flight_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + fore_length + (1 | species)")

getPlotLM(wings_yellow,model_type = "lmer",response="yellow",formula= "wing_type + Sex + temp_indv + temp_mean_sp + se_mean_sp + (1 | species)")

getPlotLM(wings_brown,model_type = "lmer",response="bb_total",formula= "flight_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + fore_length")

# getPlotLM(wings_phy_trim,phy=wings_phy,model_type = "pglmm",response="black",formula="flight_type_rm_inter + wing_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length + (1 | species__)")
# 
# getPlotLM(wings,tree_location = "../../data/odonata.tre",model_type = "pglmm",response="black",formula="flight_type_rm_inter + wing_type + Sex + temp_sd_sp + temp_q95_sp + se_indv + prec_indv + prec_mean_sp + fore_length + (1 | species__)")

```

Males have more black pigments, but less brown pigments

Hindwings have less black pigment

## Q2: A priori, do perchers have more pigments, how does sex interact, and do perchers in colder environments have more pigments and especially pigments at the wing base?

Perchers in colder environments should have more pigment at base of wings

Perchers in warmer environments should have pigments anywhere

If this is true, PCs which upweight base pigments should be correlated with flight style

```{r pattern_pcs, results='hide'}

makePatternPCA(wings,pattern_dir = "D:/wing-color/data/patterns/grouped/size_normalized",
               color_by_flight_style = TRUE,wing_type="fore")
makePatternPCA(wings,pattern_dir = "D:/wing-color/data/patterns/grouped/size_normalized",
               color_by_flight_style = TRUE,wing_type="hind")

```

Run interaction models

```{r interaction_lms, results='hide'}

# plot two way interaction
m <- getPlotLM(wings,model_type = "lmer",response="black",formula="(flight_type + Sex + temp_indv)^2 + (1 | species)",
                plotmodel_terms = c("flight_type","Sex","temp_indv"))

# plot 3 way interaction
m <- getPlotLM(wings,model_type = "lmer",response="black",formula="flight_type * Sex * temp_indv + (1 | species)",
                plotmodel_terms = c("flight_type","Sex","temp_indv"))
plot_model(m,type="pred",terms=c("flight_type","Sex","temp_indv"))


# try with wing type and fore length
m <- getPlotLM(wings,model_type = "lmer",response="black",formula="(flight_type + Sex + temp_indv + wing_type + fore_length)^2 + (1 | species)")

m <- getPlotLM(wings,model_type = "lmer",response="black",formula="flight_type * Sex * temp_indv * fore_length + (1 | species)")
plot_model(m,type="pred",terms=c("flight_type","Sex","temp_indv","fore_length"))


# try with brown
m <- getPlotLM(wings,model_type = "lmer",response="bb_total",formula="(flight_type + Sex + temp_indv + wing_type + fore_length)^2 + (1 | species)")

m <- getPlotLM(wings,model_type = "lmer",response="bb_total",formula="flight_type * Sex * temp_indv * fore_length + (1 | species)")
plot_model(m,type="pred",terms=c("flight_type","Sex","temp_indv","fore_length"))


# try pglmm

m <- getPlotLM(wings_phy_trim,phy=wings_phy,model_type = "pglmm",response="black",formula="flight_type_rm_inter * Sex * temp_indv * fore_length + (1 | species__)")



```

## Q3: Can we predict percher/flyer or m/f via logistic regression of color or pattern?

```{r logregs, results='hide'}

getPlotLogReg(wings_sp_sex,response="Sex",formula="black + brown + yellow + all_fore_pc1 + all_fore_pc2 + all_hind_pc1 + all_hind_pc2")

getPlotLogReg(wings_sp_sex,response="flight_type",formula="black + brown + yellow + all_fore_pc1 + all_fore_pc2 + all_hind_pc1 + all_hind_pc2")

plot()
```

Sex actually is predictable! May suggest that males and females have different signals or thermoregulatory functions

But flight type isn't at all!

## Q4: Do pigments or patterns have a phylogenetic signature?

```{r physig, results='hide'}

assessTraitsOnTree(wings_sp,trait_colname="black",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp,trait_colname="brown",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp,trait_colname="yellow",tree_location="../../data/odonata.tre")

# combinations of pigments 
wings_sp$bb_total <- wings_sp$black + wings_sp$brown
wings_sp$by_total <- wings_sp$yellow + wings_sp$brown
assessTraitsOnTree(wings_sp,trait_colname="bb_total",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp,trait_colname="bby_total",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp,trait_colname="by_total",tree_location="../../data/odonata.tre")


wings_sp_na0 <- wings_sp %>% 
  mutate(all_fore_pc1 = coalesce(all_fore_pc1, 0),
         all_fore_pc2 = coalesce(all_fore_pc2, 0),
         all_hind_pc1 = coalesce(all_hind_pc1, 0),
         all_hind_pc2 = coalesce(all_hind_pc2, 0))

# PCs
assessTraitsOnTree(wings_sp_na0,trait_colname="all_fore_pc1",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp_na0,trait_colname="all_fore_pc2",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp_na0,trait_colname="all_hind_pc1",tree_location="../../data/odonata.tre")
assessTraitsOnTree(wings_sp_na0,trait_colname="all_hind_pc2",tree_location="../../data/odonata.tre")

```

All responses have phylogenetic signatures

Of the individual pigments, black has the strongest followed by yellow followed by brown

Of the combined pigments, black + brown + yellow has the strongest followed by brown + yellow followed by black + brown

In other words, these pigments often occur together phylogenetically

Only the 2nd PCs have phylogenetic signature, and they are much smaller than the pigments

Rerun prep and see if NAs still stand

## Q5: Are pigments and patterns phylogenetically correlated?

```{r phylocor, results='hide'}

# brown and yellow together
assessTraitsOnTree(wings_sp,trait_colname="yellow",trait_colname2="brown",tree_location="../../data/odonata.tre")

# black and brown together
assessTraitsOnTree(wings_sp,trait_colname="black",trait_colname2="brown",tree_location="../../data/odonata.tre")


```

Both yellow-brown and black-brown are correlated

## Q6: Are pigments or patterns phylo-correlated with flight type?

```{r phylocorflight, results='hide'}

wings_sp2 <- wings_sp
wings_sp2$flight_type <- as.character(wings_sp2$flight_type)
wings_sp2$flight_type[is.na(wings_sp2$flight_type)] <- "intermediate"

# black and flight type 
assessTraitsOnTree(wings_sp2,trait_colname="black",trait_colname2="flight_type",tree_location="../../data/odonata.tre",legend=F)

# black and flight type 
assessTraitsOnTree(wings_sp2,trait_colname="yellow",trait_colname2="flight_type",tree_location="../../data/odonata.tre",legend=F)


```

Pretty much only perchers have evolved dark pigments

## Q7: Rob's flight style \* Sex interaction analysis

```{r rob, results='hide'}

library(phyr)
library(lme4)
library(sjPlot)
library(performance)
library(dplyr)
library(glmmTMB)
library(tidyr)

# relevant responses - black, brown, yellow, bb_total (black and brown total),
#   bby_total (black brown and yellow total), PC1_foreall, PC2_forall, PC1_hindall, PC2_hindall
# relevant predictors - flight_type, Sex, various temp, seasonality, precip predictors
# example lmm
# example lmm
wings$temp_indv_sc <- scale(wings$temp_indv)

wings2 <- wings %>% drop_na(Sex,temp_indv_sc, flight_type, black)
m <- lmer(black ~ flight_type * Sex * temp_indv_sc + (1 | species),
          data=wings2)
m3 <- glmmTMB(black ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
          data=wings2, family='ordbeta')
wings4 <- bind_rows(
  wings2 %>%
    group_by(flight_type) %>%
    filter(flight_type=="flyer"),
  wings2 %>%
    group_by(Sex) %>%
    filter(flight_type=="percher") %>%
    slice_sample(n = 1000))
m4 <- glmmTMB(black ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
              data=wings4, family='ordbeta')
plot_model(m4,type="pred",terms=c("flight_type","Sex","temp_indv_sc"))
# example pglmm - hard to fit bc pglmm with ordinal beta is not implemented
m <- pglmm(black ~ flight_type * Sex * temp_indv + (1 | species__),
    data = wings_phy_trim, cov_ranef = list(species=phy_for_wings), bayes=T)



wings2 <- wings %>% drop_na(Sex,temp_indv_sc, flight_type, brown)
m <- lmer(brown ~ flight_type * Sex * temp_indv_sc + (1 | species),
          data=wings2)
m3 <- glmmTMB(brown ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
          data=wings2, family='ordbeta')
wings4 <- bind_rows(
  wings2 %>%
    group_by(flight_type) %>%
    filter(flight_type=="flyer"),
  wings2 %>%
    group_by(Sex) %>%
    filter(flight_type=="percher") %>%
    slice_sample(n = 1000))
m4 <- glmmTMB(brown ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
              data=wings4, family='ordbeta')
plot_model(m4,type="pred",terms=c("flight_type","Sex","temp_indv_sc"))
# example pglmm - hard to fit bc pglmm with ordinal beta is not implemented
m <- pglmm(brown ~ flight_type * Sex * temp_indv + (1 | species__),
    data = wings_phy_trim, cov_ranef = list(species=phy_for_wings), bayes=T)


wings2 <- wings %>% drop_na(Sex,temp_indv_sc, flight_type, yellow)
m <- lmer(yellow ~ flight_type * Sex * temp_indv_sc + (1 | species),
          data=wings2)
m3 <- glmmTMB(yellow ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
          data=wings2, family='ordbeta')
wings4 <- bind_rows(
  wings2 %>%
    group_by(flight_type) %>%
    filter(flight_type=="flyer"),
  wings2 %>%
    group_by(Sex) %>%
    filter(flight_type=="percher") %>%
    slice_sample(n = 1000))
m4 <- glmmTMB(yellow ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
              data=wings4, family='ordbeta')
plot_model(m4,type="pred",terms=c("flight_type","Sex","temp_indv_sc"))



wings2 <- wings %>% drop_na(Sex,temp_indv_sc, flight_type, bb_total)
m <- lmer(bb_total ~ flight_type * Sex * temp_indv_sc + (1 | species),
          data=wings2)
m3 <- glmmTMB(bb_total ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
          data=wings2, family='ordbeta')
wings4 <- bind_rows(
  wings2 %>%
    group_by(flight_type) %>%
    filter(flight_type=="flyer"),
  wings2 %>%
    group_by(Sex) %>%
    filter(flight_type=="percher") %>%
    slice_sample(n = 1000))
m4 <- glmmTMB(bb_total ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
              data=wings4, family='ordbeta')
summary(m4)
r2_nakagawa(m4)
plot_model(m4,type="pred",terms=c("flight_type","Sex","temp_indv_sc"))


wings_black$temp_indv_sc <- scale(wings_black$temp_indv)

wings2 <- wings_black %>% drop_na(Sex,temp_indv_sc, flight_type, black)
m <- lmer(black ~ flight_type * Sex * temp_indv_sc + (1 | species),
          data=wings2)
m3 <- glmmTMB(black ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
          data=wings2, family='ordbeta')

m4 <- glmmTMB(black ~ (flight_type * Sex) + temp_indv_sc + (1 | species),
              data=wings4, family='ordbeta')
plot_model(m3,type="pred",terms=c("flight_type","Sex","temp_indv_sc"))
# example pglmm - hard to fit bc pglmm with ordinal beta is not implemented
m <- pglmm(black ~ flight_type * Sex * temp_indv + (1 | species__),
    data = wings_phy_trim, cov_ranef = list(species=phy_for_wings), bayes=T)

```

Pretty much only perchers have evolved dark pigments

## Q8: Rob's black / black + brown + yellow analysis

This metric is high if there is more black than other colors

If all the pigments in a wing are black, the value is 1.

If there are half brown and half black, the value is 0.5.

If there is no black, the value is 0

Compared to the black variable, which is 0 if there is no black, 0.5 if there is half black and any amount of brown

```{r rob_prop_black, results='hide'}

wings2 <- wings_black
wings2$Area_total_sc <- scale(wings2$Area.FW..mm2. + wings2$Area.HW..mm2.)
wings2$length_sc <- scale(wings2$Length..inner..FW..mm.)

wings2$prop_black <- wings2$black / wings2$bby_total
m2_propbl <- glmmTMB(black ~ flight_type * length_sc + Sex + (1 | species),data=wings2,family='ordbeta',na.action="na.omit")
summary(m2_propbl)
plot_model(m2_propbl,type="pred",terms=c("flight_type","Sex","length_sc"))

m2_propbl <- glmmTMB(prop_black ~ flight_type * length_sc + Sex + (1 | species),data=wings2,family='ordbeta',na.action="na.omit")
summary(m2_propbl)
plot_model(m2_propbl,type="pred",terms=c("flight_type","Sex","length_sc"))

test <- dplyr::filter(wings,wing_type == "fore")
test$Area.HW..mm2.
```

Run interaction models

## Q8: Rob's zero inflated models

```{r zi, results='hide'}
library(glmmTMB)
library(dplyr)
library(tidyr)

wings$Area_total <- wings$Area.FW..mm2. + wings$Area.HW..mm2.

wings2 <- wings %>% drop_na(Sex,temp_indv, flight_type, black, Area_total, bby_total)
wings2$temp_indv_sc <- scale(wings2$temp_indv)
wings2$Area_total_sc <- scale(wings2$Area_total)
wings2$black01 <- ifelse(wings2$black  > 0, 1, wings2$black)

library(lme4)
area_temp_sex <- glmmTMB(black ~ Area_total_sc * temp_indv_sc + Sex  + (1 | species), data=wings2, family=beta_family(link="logit"),  ziformula = ~., na.action = "na.fail")

area_temp_sex_glm <- glmer(black01 ~ Area_total_sc * temp_indv_sc + Sex  + (1 | species), data=wings2,
                         family="binomial")
plot_model(area_temp_sex_glm, type="pred", c("Area_total_sc","temp_indv_sc","Sex"))


sex_flight_temp <- glmmTMB(black ~ Sex * flight_type + temp_indv_sc  + (1 | species), data=wings2, family=beta_family(link="logit"),  ziformula = ~., na.action = "na.fail")
summary(sex_flight_temp)
sex_flight_temp_log <- glmer(black01 ~ Sex * flight_type + temp_indv_sc  + (1 | species), data=wings2,
                         family="binomial")
summary(sex_flight_temp_log)
plot_model(sex_flight_temp_log, type="pred", c("flight_type","temp_indv_sc","Sex"))

library(performance)
r2_nakagawa(sex_flight_temp_log)
r2_nakagawa(area_temp_sex_glm)
# strong set of effects tied to area '

# plot logistic 

# plot conditional
library(marginaleffects)
plot_predictions(area_temp_sex, type="conditional", condition = list( "Area_total_sc","temp_indv_sc", "Sex"))
plot_predictions(sex_flight_temp, type="conditional", condition = list("flight_type", "Sex","temp_indv_sc"))

library(sjPlot)
plot_model(area_temp_sex, type="pred", c("Area_total_sc","temp_indv_sc", "Sex"))
plot_model(sex_flight_temp, type="pred", c("flight_type", "Sex","temp_indv_sc"))

# logistic regressions 
```

Run interaction models

    {r zi, results='hide'}

    m2_totbl_b <- glmmTMB(black ~ Area_total_sc * temp_indv_sc + Sex  + (1 | species), data=wings2, family=beta_family(link="logit"),  ziformula = ~., na.action = "na.fail")
